---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r logo, echo=FALSE, out.width="600px", fig.align="center"}
knitr::include_graphics(
  "/Users/aloza/OneDrive/Desktop/Bayou/STAT-306-Logo.png"
)
```

# Welcome To The Bayou!

<!-- badges: start -->
<!-- badges: end -->

The name of this package is “Bayou”. Now you may be wondering why someone would name a package like that. This Package has been built around my favorite movie, The Princess and the Frog. This movie is very near and dear to my heart, and even though I am getting older, the movie still holds the number 1 spot for my favorite movie.  One of the main messages from the movie is that hard work and hope are not mutually exclusive. You need both hand-in-hand to work together in order to achieve something that you really want in your heart. 

## Installation

You can install the development version of Bayou from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("NicoleLincoln/Bayou")
```

## Accessibility: 

The theme of this package consists mainly of very Earthy colors, a lot of greens, neutrals, and blues, all of which align with the overall color palette of the movie. These colors are very saturated and pass the color contrast test so people can differentiate between them. The font for the theme is also a Serif font, so it is easier for people to read, for those who may have reading difficulties, at an adequate size for trading purposes. Having knowledge of accessibility is important in terms of data science because we want our visualization to be readable to as many people as we can to help people learn from the information/messages that we have uncovered. 
	
## Helper Function: 

Now, I am in a Biology lab where we research Fireflies. While we are collecting fireflies, we have a very specific way of documenting our fireflies/. We give them an identification number, write down the collection time, mass, temperature, and humidity levels. However, one thing that we use is a metric called season day. Essentially, the first time we see a firefly and collect it is considered day one. So, this gave me the idea: what if there was a function that automatically arranged a data set in ascending order and assigned the number 1 to the number of total rows in the data set. This would allow us to assign an ordered value associated with a specific individual. Having individuals associated with their ordered value can allow us to make visualizations that have a new ordered value a possible x or y variable. 
	
## Example

This is a basic example of how to use the Helper Function. Please refer to the vignette titled 'BayouDataMoves'to learn more bout what this does:

```{r example}
library(Bayou)

castle <- readr::read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2026/2026-09-01/world_castles.csv",
  n_max = 10,
  show_col_types = FALSE
)
Trumpets(castle, "year")
```

# Theme: 

The theme of this Package includes a combination of our accessibility principles and customization colors. I designed the theme to be as customization as possible, but still keeping a rather minimalist look with some pops of color. I think for a lot of data visualizations, the addition of too many add-ons and possibly even data points can make the visualization look too cluttered and make the visualization harder to read. This is the opposite of what we want from a visualization. You want your data to speak for itself, for your readers to be able to understand the message that you are trying to put forth. So I wanted to create a theme that makes the visualization minimalist, but still have some visual appeal. 


## Vignette: 

I have created two different vignettes for this package to explain in detail two different sections of my package. The first one is labeled ‘BayouDataMoves’. This vignette explains in detail different data moves that can be performed on an example data set named ‘castle’. This vignette also includes an example of the use of my helper function. It is important to read in detail what these data moves and the helper function do because it can make the data exploration process easier. I chose to include this because these data moves have been really helpful to me when learning how to explore different data sets to understand patterns and answer certain questions. The second vignette is named ‘Theme_Evangeline’; this goes into detail about the background of the theme and how I found it to be useful for myself and how I think it could possibly help others in their data communication process. 

## Colors 

I have created multiple different color combinations for people to use to add to the theme. They all surround the theme of The Princess and the Frog while all keeping neutral colors with some hints of yellow and purple. Please run the color combinations before creating any visualizations so that when you scale-fill or scale-color, they are already loaded into the environment. I wanted people to have the opportunity to choose which combination they wanted to use because there is not one single color combination that represents The Princess and the Frog, but it could be multiple. 

