---
title: "BayouDataMoves"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BayouDataMoves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Loading the Package

```{r setup}
library(Bayou)
```

```{r}
library(tidyverse) 
castle <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2026/2026-09-01/world_castles.csv")
```

## Data Moves

*Filter*

  Sometimes, looking at large data can be very overwhelming. A way to subset the data to only include a subset of variables that we care about. To do that, we can filter a column or row to only certain values. 

	For example, let’s take our castle data set. In the category column, there are four different landmark categories that the individuals fall into. Now, in the Princess and the Frog Princess and the Frog, it is said that Tiana and Naveen lived in a castle, living happily ever after. Keeping this in mind, let's see what our data look like when we filter the data to only the landmarks whose category is a castle. 

```{r}
castle %>%
  filter(category == "castle") %>%
  summarise(castle_sum =n())
```
  Now, this code uses a ‘%>%’. What this does is essentially follow sequential steps. For instance, in this code chunk we start with the data set castle, then we filter that data set to only those whose category is “castle”. Then, from the filtered data set, we summarized the number of landmarks that are castles by counting the number of rows that are in the filtered data set. 

So, in the large ‘castle’ data set, there are 2591 landmarks that are castles. 


*Group By/Summarise*

  Now, there are other ways to organize our data to make it less overwhelming. We can also perform a Group By/Summarize procedure. What this means is that they essentially group the data set by a certain variable, typically a column. 

	Now, wouldn’t it be nice to see how many landmarks were built in each century? From this, we could learn which century had the most landmarks, which the least, and even rises and falls throughout time. Now let us run the code to Group the data set and summarize the results.

```{r}
castle %>%
  filter(category == "castle")%>%
  group_by(century) %>% 
  summarise(cent_sum = n())

castle %>%
  filter(category == "castle")%>%
  group_by(century) %>% 
  summarise(cent_sum = n())%>%
  arrange(century)
```
Now, following through the code, we see the filtered data move again (if reference needed, please read the above section titled: Filter). We have subset our data to only the landmarks that are categorized as a castle. Then, from that data set, we are grouping the castles by the century the founding year falls in. Similar to above, we are summarizing the data by presenting the number of rows that belong to each century of the year the castle was founded. 

	Based on the summarized data, the century with the highest number of castles founded is the 13th century, with 279 castles 

	In the second code chunk, you can see that there was a new line of code added into a new pipe: ‘arrange(century).’ To go more into detail on what this line of code does, please read the section titled: Arrange. 

From this new code chunk, we learn that there is an increase in the number of castles being founded from the 20th century to the 13th century, followed by an almost steady decline following the 13th century. 


*Join*

  Now, the ‘join’ data move allows you to join two data sets based on an identical column, unifying them into one. This can help us expand the dataset so we can ask more questions about the topic. 

  To do this, I have created a new data set that is a subset of the original that has been manipulated and joined back to the original data set. 

```{r}
totals <- castle %>%
  group_by(category) %>%
  summarise(total = n())

totals <- totals %>%
  mutate(prop = total / 5793)

castle <- castle %>%
  left_join(totals, by = (c="category"))
  
```
  The first two chunks of code deal with the creation and manipulation of the new data set titled ‘totals’.  The first chunk creates the new data set, ‘totals’, which is the castle data frame grouped by category (similar to the Filter section) and only contains the total number of rows that are in each category. So this results in a 4 (category) by 2 table in which the castle has a total of 2591, the fortress has 1180, the palace has 934, and the ruin has 1088. The second code chunk mutates (see the section on Mutate for more information) the new data set ‘totals’ to now have a new column named “prop” which takes the totals of each category and vided is by 5793, which is the total number of landmarks in the data set, creating a proportion for each category. 

	The final code chunk is responsible for the actual join component. Within the castle data set, we are writing over the original ‘castle’ data set with another called castle. The new ‘castle’ data set is comprised of joining the old castle data set with the one named ‘totals’ by the unifying column “category. Both data sets share a column named “category” with the same possible values, so RStudio is able to combine both of  the data sets. Now if you run the code you should see the ‘castle’ data set now has two addition columns the are labeled “total” and “prop”. This should look familiar to the original ‘totals’ data set. What the code did was essentially or every row whose category was "castle" the code added the value of the total for castle and the corresponding prop value and so on and so forth for the other three categories.  Now we can compare the proportion of each category with other variables in the data set, something we were not able to do before. 


*Arrange*

  This data set has a unique column named “fame_rank”. “Fame_rank” is the global fame rank. Now, what is the purpose of a rank if you cannot tell who is number 1? This can be easily achieved by simply arranging our data set. Now, for this data move, it is important to know the difference between ascending and descending. 

Descending is going from highest to lowest. 

Ascending is going from lowest to highest. 

  Remember to think A comes before D, so ascending is the one where the smallest comes first. 


```{r}
castle %>%
  arrange(fame_rank)

castle %>%
  arrange(desc(fame_rank))
```
  Now this code arranges the data set in ascending order based on the variable fame-rank. Now, in the arrange function, ascending is the default. Therefore, nothing else is needed for the code. In the second chunk of code, the fame-rank is organized in descending order. You can see the data set remains the same; only the order is changing. 


*Mutate*

  Mutations in a modern sense are typically tied to biology and negative connotations.  However, in RStudio, a mutation means that you are adding a new column in your data set. Now, in this data set, because we have a wide range of dates when the landmarks were founded, I wanted to know which landmarks are a part of ancient history and which are considered to be modern history. Google has a consensus that anything before the year 500 is considered Ancient History, while everything after the year 500 is considered Modern History. So, the following code does just that. 


```{r}
castle %>%
  mutate(history = ifelse(castle$year < "500", "Ancient", "Modern"))

```
  Based on the code, the castle dataset is mutated to create a new column named “history” in which, if the year of the landmark is below 500, then the column would be filled with the word “Ancient,” but if the year doesn’t pass this requirement, then the column would instead have the word “Modern”. This check is done for every case and landmark in the dataset until the whole column is filled. 

# Helper Function

  One of the main goals of this package is to include a useful Helper Function. This placed a lot of pressure on me because there was pressure to create a Function that would be the most efficient and universal Helper Function for people to use. However, a Helpful function is a helper function even if it only helps a subset of people. 

	I am in a Biology lab where we research Fireflies. While we are collecting fireflies, we have a very specific way of documenting our fireflies/. We give them an identification number, write down the collection time, mass, temperature, and humidity levels. However, one thing that we use is a metric called season day. Essentially, the first time we see a firefly and collect it is considered day one. So, this gave me the idea: what if there was a function that automatically arranged a data set in ascending order and assigned the number 1 to the number of total rows in the data set. This would allow us to assign an ordered value associated with a specific individual. 

	Now, this is essentially what I created my function to do. The function begins by taking the two inputs. The first one is the data frame, and the second is the column name. The function will then arrange the data set, followed by a mutation that creates a new column named “Order-id,” which is the value of the row number after the arrangement. What you will get in return is a new data set named Trumpets with the changes made to the original. 

```{r}
Trumpets(castle, "year")
```

  I wanted to create a function that creates a new data set yet keeps the old one so that if anything happens, you can always refer back to the original. This was very important to me because I have run into the issue where I have made too many changes to the data set that sometimes RStudio will not be able to perform certain functions, so being able to go back to the original and reference it is a good way to retrace your steps. 
